### 2022-04-25 paper/main.tex
added python package, paper outline

\documentclass[11pt]{extarticle}

\title{Assessing the Performance of Matching Methods in Observational Studies}
\author{Jack Potrykus}
\date{\today}

\input{preamble.tex}

\begin{document}

\maketitle

\section*{Abstract}
\tableofcontents
\newpage

\section{Introduction}

\section{Literature Review}

\subsection{Measuring Similarity}

\subsubsection{Propensity Scoring}

Rosenbaum and Rubin, 1983
\begin{itemize}
  \item The \emph{propensity score} is most often accredited to \ldots
  \item They offer a more general definition of a score: anything affording conditional independence
  \item Score seeks to replicated randomized trial
  \item Review of properties (particularly: when can we make unbiased estimates of ATE)
  \item Proposed use cases
\end{itemize}

Methods for calculating propensity scores: Garrido et al, 2014
Note: Blocking-based balance metrics

\paragraph{Extensions}
\begin{itemize}
  \item (Optimal) Caliper width: Austin 2011
  \item Prognostic score: Hansen
  \item Miettinen score is the root of the above
  \item Joint use of Prognostic, Propensity, + Mahalanobis, Leacy and Stuart 2014
  \item Imai and Van Dyk (make sure to include)
\end{itemize}

\subsubsection{(Coarsened) Exact Matching}

\begin{itemize}
  \item Iacus King Porro, 2011: MIB methods
  \item Iacus King Porro, 2012: Causal inference without balance checking
\end{itemize}

\subsection{Balance Assessment}

\subsection{Matching Algorithms}

\section{Experiments}

\subsection{Data Generation}

Papers with data generation:
\begin{itemize}
  \item Austin 2011 (optimal caliper widths)
  \item Stuart et al: 2018
\end{itemize}

\section{Results}

\section{Discussion}

\section{Conclusion}

\end{document}