### 2022-04-26 paper/main.tex
outlined most of lit review

\documentclass[11pt]{extarticle}

\title{Assessing the Performance of Matching Methods in Observational Studies}
\author{Jack Potrykus}
\date{\today}

\input{preamble.tex}

\begin{document}

\maketitle{}

\section*{Abstract}
\tableofcontents{}
\newpage{}

\section{Introduction}
\label{sec:Introduction}

In the context of observational data, a bipartite matching problem is minimally defined by:
\begin{itemize}
  \item \mat{X}, an $n \times p$ matrix of feature values;
  \item \vec{z}, an $n$-vector of binary treatment assignments;
  \item $d : \reals^p \times \reals^p \mapsto \reals^+$, a function which computes some ``distance'' or ``cost of matching'' between any two row vectors \vec{x_i} and \vec{x_j} of \mat{X}.
\end{itemize}
% The output

\set{S}, \vec{b}, \mat{A}, \vec{\upbeta}
test \cite{iacus_multivariate_2011}

\section{Literature Review}

\subsection{Measuring Similarity}
Iacus, King, Porro 2011 - set the scene

\subsubsection{Propensity Scoring}

Rosenbaum and Rubin, 1983
\begin{itemize}
  \item The \emph{propensity score} is most often accredited to \ldots
  \item They offer a more general definition of a score: anything affording conditional independence
  \item Score seeks to replicated randomized trial
  \item Review of properties (particularly: when can we make unbiased estimates of ATE)
  \item Proposed use cases
\end{itemize}

Methods for calculating propensity scores: \cite{garrido_methods_2014}
Note: Blocking-based balance metrics

\paragraph{Extensions}
\begin{itemize}
  \item (Optimal) Caliper width: \cite{austin_optimal_2011}
  \item Prognostic score: \cite{hansen_prognostic_2008}
  \item Miettinen score is the root of the above \cite{miettinen_stratification_1976}
  \item Joint use of Prognostic, Propensity, + Mahalanobis, \cite{leacy_joint_2014}
  \item \cite{imai_causal_2004}
\end{itemize}

\subsubsection{(Coarsened) Exact Matching}

\begin{itemize}
  \item Iacus King Porro, 2011: MIB methods
  \item Iacus King Porro, 2012: Causal inference without balance checking
\end{itemize}

\subsection{Balance Assessment}

Garrido et al, 2014: don't use the outcome in the matching
Garrido et al, 2014: balance in mean does not imply balance in scores

\subsection{Matching Algorithms}

\parencite{rosenbaum_optimal_1989}
\parencite{ho_matchit_2011} (Greedy)
\parencite{khan_efficient_2016} (hungarian)
\parencite{munkres_algorithms_1957} (hungarian)
Hungarian -- Munkres, Khan et al 2016

\section{Python Package}

\section{Experiments}

\subsection{Data Generation}

Papers with data generation:
\begin{itemize}
  \item \cite{austin_optimal_2011}
  \item \cite{stuart_prognostic_2013}
\end{itemize}

\section{Results}

\section{Discussion}

\section{Conclusion}

% NOTE: SEE LINK BELOW
% https://tex.stackexchange.com/questions/8458/making-the-bibliography-appear-in-the-table-of-contents
% Special handling of bibliography awaits!!
\cleardoublepage
\phantomsection
\addcontentsline{toc}{section}{Bibliography}
\printbibliography

\end{document}